---
title: "Report for homework"
author: "Nurşen Töre"
date: "3/15/2021"
output: html_document
---
```{r libraries, echo=FALSE,message=FALSE}
library(tidyr)
library(magrittr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(readr)
```

# Loading and preprocessing the data

```{r loading}
fileurl="https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileurl,'./activity.zip', mode = 'wb')
unzip("activity.zip", exdir = getwd())
dat<-read_csv("activity.csv")
head(dat)
```

# What is mean total number of steps taken per day?

Calculate the total number of steps taken per day

```{r totwalk}
totwalk<-dat%>%
  group_by(date)%>%
  summarize(totwalk=sum(steps,na.rm=TRUE))
```

Make a histogram of the total number of steps taken each day

```{r histogram of totwalk}
with(totwalk,hist(totwalk,main=" Total Walk Histogram",
                  xlab="Total Steps",col="darkmagenta"))
```

Calculate and report the mean and median of the total number of steps taken per day

```{r mean and median of totwalk}
summ<-summary(totwalk$totwalk)
summ["Median"]
summ["Mean"]
```
# What is the average daily activity pattern?

Make a time series plot  of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r time series average steps}
avewalk<-dat%>%
  group_by(interval)%>%
  summarize(avesteps=mean(steps,na.rm=TRUE))
with(avewalk,plot(interval,avesteps,type = "l",main="Average steps throughout day"))

```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r max ave steps}
avewalk%>%filter(avesteps==max(avesteps))%>%
  select(interval)
```

# Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r NAs}
sum(is.na(dat))
```
Devise a strategy for filling in all of the missing values in the dataset.
I will use the mean for that 5-minute interval.

Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r remove nas}
dat<-dat%>%group_by(interval)%>%
  mutate(steps = replace_na(steps, mean(steps,na.rm=TRUE)))%>%
  ungroup()
```
Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r histogram for imputed table}
totwalk<-dat%>%
  group_by(date)%>%
  summarize(totwalk=sum(steps,na.rm=TRUE))
summ<-summary(totwalk$totwalk)
summ["Median"]
summ["Mean"]
with(totwalk,hist(totwalk,main=" Total Walk Histogram",xlab="Total Steps",col="darkmagenta"))

```

# Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r create factor variable}
weekavesteps<-dat%>%mutate(isweekday = case_when(
    weekdays(date)=="Monday"  ~ "Weekday",
    weekdays(date)=="Tuesday"  ~ "Weekday",
    weekdays(date)=="Wednesday"  ~ "Weekday",
    weekdays(date)=="Thursday"  ~ "Weekday",
    weekdays(date)=="Friday"  ~ "Weekday",
    weekdays(date)=="Saturday"  ~ "Weekend",
    weekdays(date)=="Sunday"  ~ "Weekend"),
    isweekday = factor(isweekday, levels=c("Weekday","Weekend")))%>%
  group_by(isweekday,interval)%>%
  summarize(avesteps=mean(steps,na.rm=TRUE))%>%ungroup()
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r create panel plot}
p <- ggplot(weekavesteps, aes(interval, avesteps)) 
p+ geom_line()+facet_wrap(~ isweekday)
```
